---
title: "cleaning dataset"
author: "Marina Sánchez Moreno"
date: "2025-03-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r libraries, results='hide'}
library(knitr)
library(readr)
library(dplyr)
library(plyr)
library(tidyr)
library(stringr)
library(tidyverse)
```

## Introducing the dataset

The dataset we will work on has been built by partners in T2.4 - "Learning materials
quality assurance and certification framework" of the European project [Skills4EOSC](https://www.skills4eosc.eu/), reviewed and published together with [D2.7 Community-endorsed quality assurance and certification framework for professional
training and qualifications - final version](https://zenodo.org/records/15731878) as one of the outputs from project. 

The main dataset used to build the questionnaire includes the quality assurance (QA) 
indicators that will measure the compliance of open science courses and modules to 
the Skills4EOSC QA and Certification Framework. Moreover, a small set of questions 
was added to collect relevant information about the user for future analysis on 
responses and feedback to improve the app.

The first dataset (Quality Assurance questions) contains the following information:

  - *id_ind*:	unique identification number for each indicator, regardless their section
  of category. Letter of subframework followed by a number
  - *dependence*:	It provides the id of the indicator it is subordinated to. It 
  determines whether
  an indicator is subordinated to another indicator's response. It reduces redundant
  or unecessary questions. For instance, if the user has indicated that their learning
  resource does not provide a title, the app will not ask more questions about the title.
  - *dependence_value*:	when there is dependence (dependence is not NA), this variable
  captures the needed answer choice in the parent question. For instance, if E02 is dependent
  of E01.
  _dependence_ for E02 would be "E01" and _dependence_value_ would be the answered 
  in E01 that activates question E02 (e.g.: "Yes")
  - *question_txt*:	The user will need to read and answer these questions to assess
  QA compliance. Equivalent to "Checklist question" in the general implementation v3
  document  
  - *subframework*:	It shows the subframework the indicator is part of. This variable
  is just for partners to easily update the database
  - *property*:	The topic/item the indicator refers to (one word). It is equivalent 
  to the column "property" in the general implementation v3 document
  - *category*:	It is the category that the property of the indicator is part of. 
  It helps organizing the indicators in a user-friendly way
  - *section*: It is the section that the category of the indicator is part of. It 
  helps organizing the indicators in a user-friendly way
  - *minimal_detailed*: 0 if the indicator is detailed and 1 if it is minimal (old 
  "essential")
  - *answer_user*: it tracks the user answer (YES=1 or NO=0)
  - *bestpractices*	Plain text with the best practices associated to the indicator 
  - *notes*: additional information (e.g. definitions) provided to the user regarding
  the question. Generally, it provides definitions about a term.
  - *keywords_ind*:	list of terms/topics related to the indicator. 
    
The second dataset regarding personal data questions follows a similar structure.
Therefore, no much pre-processing will be needed.

Note: In the raw dataset for the self-assessment (qa_data) there are some links in **notes**,
**bestpractices** and **question**. To be able to render this links in the app, links are
formatted as html link tags:

  - <a href="https..." target="_blank">clickable text</a>

Then, in the app script, this columns are read under the **HTML()** function.


## Pre-processing the dataset

### Quality Assurance dataset

- Read and clean raw data

```{r}

raw_data1 <- read_csv2 (".//www/qa_data.csv", col_names = TRUE, na= "" )

qa_data <- raw_data1 %>% 
  # Remove empty rows
  filter(!is.na(question_txt)) 

```

New variables that we need to create:


  - *options*: List of possible responses (choices) to the question, separated by ";". 
    - For most of our questions (yes/no questions) this will be "Yes;No". For a list of,
    this will be a mc type (multiple choice), that is, "Yes;No;Not applicable"
  - *required*: logical TRUE/FALSE signifying if a question is required. 
    - We will set this as TRUE for minimal indicators only and FALSE for detailed ones.

```{r}

# CREATE COLUMNS
# Changes to columns
qa_data <- qa_data %>% 
  # Rename question_txt with question
  dplyr::rename("question"="question_txt") %>% 
  # Rename id_ind to input_id
  dplyr::rename("input_id" = "id_ind") %>% 
  # Create a column "options" with "Yes;No" or "Yes;No;Not applicable"
  dplyr::mutate(options = case_when(
    input_id == "G15"  ~ "Yes;No;Not applicable",
    input_id == "M06"  ~ "Yes;No;Not applicable",
    input_id =="M07" ~ "Yes;No;Not applicable",
    input_id == "F07" ~ "Yes;No;Not applicable",
    input_id == "E04" ~ "Yes;No;Not applicable",
    TRUE ~ "Yes;No" )) %>% 
  #Based on the values from "options", create a column input_type with "y/n" or "mc"
  dplyr::mutate(input_type= "mc") %>%
  # Create required column (all questions are set to FALSE in this version)
  dplyr::mutate(required= FALSE) %>%
  # Previous version: Create column required with TRUE when the indicator is minimal and FALSE ifelse
  #dplyr::mutate(required= case_when(minimal_detailed =="minimal"~ TRUE,
  #                                  minimal_detailed !="minimal"~ FALSE))%>%
  # Create column page from column section
  dplyr::rename("pages"="section") %>% 
  select(!(starts_with("source")))

```


### Personal data 

For the dataset regarding personal data, we created the csv using the required variables,
so no great changes are needed. We only need to read the dataset, normalised it as the previous
one and merge both:

```{r}
# Read csv file
raw_data2 <- read_csv2 (".//www/personal_data.csv", col_names = TRUE, na= "" )

personal_data <- raw_data2 %>% 
  # Remove empty rows
  dplyr::filter(!is.na(question)) %>%
  # Rename section column to pages
  dplyr::rename("pages"="section") %>% 
  # Change required values from 1 to TRUE and 0 to FALSE for all questions 
  # (in this version, all questions are set to false anyway)
  dplyr::mutate(required= case_when(required =="1"~ TRUE, required!="1"~ FALSE)) %>% 
  # Based on the values from "option" by group (duplicates of the same indicator),
  # create a column "options" (values separated by ";")
  group_by(input_id) %>% 
  dplyr::reframe(options = stringr::str_c(option, collapse=";"), across()) %>% 
  # delete duplicates
  dplyr::distinct(input_id, .keep_all=TRUE) %>% 
  # delete old column option
  dplyr::select(-option)
  

```

## Merging datasets

```{r}

survey <- rbind.fill(qa_data, personal_data)

survey_clean <- survey %>%
  filter(input_type %in% c("mc", "text", "textSlider", "select")) %>%
  distinct(input_id, .keep_all = TRUE) %>%
  filter(!is.na(input_id), input_id != "") %>%
  mutate(
    options = ifelse(
      is.na(options),
      NA_character_,
      gsub("\\s*;\\s*", ";", trimws(options))
    )
  )

```


## Multipage survey

To structure the survey by page, we create a variable *page* coding the values from
variable *section*. 
  - *Note: This is not actually needed since we are not using the shinysurvey library anymore.*

Moreover, we want pages to be ordered in a certain way: 
  - 0. Personal Data
  - 1. Content & Structure
  - 2. Implementation
  - 3. Evaluation
  - 4. Licencing & Ethics

Then, we would need to order the dataframe by page, but in our case, to make sure
questions are also ordered in a coherent way within the same page and keeping dependent
questions close to their parent question, we need to order by property. By doing so, 
questions will be also ordered by category and section/page. We also remove break lines
in notes and bestpractices (otherwise, this may give you errors in the app when rendering).


```{r}

# Arrange order by page
survey_clean <- survey_clean %>% 
  dplyr::arrange(pages)

property_order <- c("About you", "Title", "Authors", "Language",
                    "Keywords", "Metadata", "Vocabularies",
                    "Catalogue", "Repository", "Audience", "Expertise level", "Goal",
                    "Objectives and outcomes", "Program", "Unit plan", "Logical unit",
                    "Learning style", "Technology", "File formats", "Prerequisites",
                    "Accessibility", "Trainer", "Instructor kit", "Timescale", "Delivery",
                    "Costs", "Access rules", "Reused resources", "Open community", "Versioning",
                    "Activities", "Grading", "Course certification", "Feedback", "Rating",
                    "Learning experience quality", "Intellectual Property Rights (IPR)",
                    "ToS and personal data", "Ethics Information")

# Arrange by property (and consequently, by category and page)
survey_clean <- survey_clean %>% 
  
  dplyr::arrange(match(property, property_order))

# Remove possible break lines to avoid errors
survey_clean$notes <-str_replace_all(survey_clean$notes, '\\n', '')
survey_clean$bestpractices <-str_replace_all(survey_clean$bestpractices, '\\n', '')


# Check remaining line breaks
#str_detect(survey_clean$notes, '\\n')
# str_detect(survey_clean$bestpractices, '\\n')
# If you run the lines above, you should get NA (when there are no notes) and/or FALSE
# When there are notes but there are no breaklines
```


This table represents the intended order and structure of the self-assessment test:

```{r hierarchy qs table, echo=FALSE, results='asis'}
table <- survey_clean %>% 
  select(pages, category, property)
kable(table, caption = "Organised Indicators Table")
```


## Save dataset

```{r}

write.csv2(survey_clean, ".//www/self-assessment_clean.csv")


```


## General feedback survey

We repeat the same process for the general feedback survey. This survey asks users
about their previous experience in quality assurance of learning materials, as well as
their perception and opinion about the Skills4EOSC Quality Assurance in general and
related tools.

The structure of this dataset is very similar to the personal_data dataset (section 
0. Background information), with questions duplicated by answers options.

```{r}
# Read csv file
raw_generalfeedback <- read_csv2 (".//www/general_feedback.csv", col_names = TRUE, na= "" )

general_feedback <- raw_generalfeedback %>% 
  # Remove empty rows
  dplyr::filter(!is.na(question)) %>%
  # Rename section column to pages
  dplyr::rename("pages"="section") %>% 
  # Change required values from 1 to TRUE and 0 to FALSE for all questions 
  # (in this version, all questions are set to false anyway)
  dplyr::mutate(required= case_when(required =="1"~ TRUE, required!="1"~ FALSE)) %>% 
  # Based on the values from "option" by group (duplicates of the same indicator),
  # create a column "options" (values separated by ";")
  group_by(input_id) %>% 
  dplyr::reframe(options = stringr::str_c(option, collapse=";"), across()) %>% 
  # delete duplicates
  dplyr::distinct(input_id, .keep_all=TRUE) %>% 
  # delete old column option
  dplyr::select(-option)

# Filter by correct input types, delete duplicates and create options column with
# possible answers by question
general_feedback <- general_feedback %>%
  filter(input_type %in% c("mc", "text", "textSlider", "select")) %>%
  # delete duplicates
  distinct(input_id, .keep_all = TRUE) %>%
  filter(!is.na(input_id), input_id != "") %>%
  mutate(
    options = ifelse(
      is.na(options),
      NA_character_,
      gsub("\\s*;\\s*", ";", trimws(options))
    )
  )

# Save dataset
write.csv2(general_feedback, ".//www/general_feedback_clean.csv")


```


## References

- 
-
-
-
-



